\chapter{IMEX Linear Multistep Methods}
For equations whose righthand side comprise of a stiff linear part and a nonstiff nonlinear component, a popular class of methods to apply are the implicit-explicit linear multistep methods (IMEX LMMs), with the most simple being an application of forward Euler to the nonlinear component and backward Euler to the linear, stiff component. 

\section{IMEX LMM Formulas}
In \cite{ascher1995implicit}, IMEX LMMs up to order four are investigated and a select number of schemes are singled out for their extensive use in the literature or for desired properties such as strong high frequency damping. As we are familiar with the first order variant, we begin by listing the second order methods of interest. These, and the higher order variants, will be presented as applied to the ODE 
\begin{align*}
u' = f + g, 
\end{align*}
where $f$ we identify as the nonlinear component and $g$ the stiff component. 
\\ \noindent
\textbf{Second order methods}\\
CNAB:
\begin{align}
\frac{u^{n+1}-u^n}{\Delta t} 
= \frac{3}{2} f^n - \frac{1}{2}f^{n-1} 
+ \frac{1}{2}(g^{n+1} + g^n), 
\label{cnab}
\end{align}\\
mCNAB:
\begin{align}
\frac{u^{n+1}-u^n}{\Delta t} 
= \frac{3}{2}f^n - \frac{1}{2} f^{n-1}
+ \frac{9}{16}g^{n+1} 
+ \frac{3}{8}g^n
+ \frac{1}{16}g^{n-1},
\label{mcnab}
\end{align}
\\
CNLF:
\begin{align}
\frac{u^{n+1}-u^{n-1}}{2\Delta t}
= f^n + \frac{1}{2}(g^{n+1} + g^{n-1}),
\label{cnlf}
\end{align} \\
SBDF2:
\begin{align}
\frac{3u^{n+1}-4u^n+u^{n-1}}{2\Delta t} 
= 2f^n - f^{n-1} + g^{n+1}.
\label{sbdf2}
\end{align}

\noindent
\textbf{Third order methods}\\
SBDF3:
\begin{align}
\frac{1}{\Delta t}\left(\frac{11}{6}u^{n+1} - 3u^n + \frac{3}{2}u^{n-1} - \frac{1}{3}u^{n-2} \right) 
= 3f^n - 3f^{n-1} + f^{n-2} + g^{n+1}.
\label{sbdf3}
\end{align}

\noindent
\textbf{Fourth order methods}\\
SBDF4:
\begin{align}
\frac{1}{\Delta t}\left(\frac{25}{12}u^{n+1} - 4u^n + 3u^{n-1} - \frac{4}{3}u^{n-2} + \frac{1}{4}u^{n-3} \right) 
= 4f^n - 6f^{n-1} + 4f^{n-2} -f^{n-3} + g^{n+1}.
\label{sbdf4}
\end{align}

Having given a collection of IMEX LMM formulas, we are now ready to apply such IMEX schemes to the modified test equation.

\section{Parameter Selection}
Let us restate here the modified test equation and some basic assumptions we make.
The modified test equation is 
\begin{align*}
        u' =  (1-p)\lambda u + p\lambda u.
\label{mte}
\end{align*}
In our analysis, we will assume $p>0$ and $\lambda < 0$. The goal is then, for each IMEX LMM we consider, to identify a constraint on the parameter $p$, such that when satisfied, we may freely choose the time step-size without being subject to a stability restriction (eg. $\Delta t < C\Delta x^2$).

\subsection{Von Neumann Polynomials}
The resulting amplification polynomials from applying an $n$th order IMEX LMM to the modified test equation will be a degree $n$ polynomial in the amplification factor, $\xi$. The tool of choicie for the analysis of these amplification polynomials is the theory of Von Neumann polynomials \cite[Chapter 4]{strikwerda2004finite}. Below we give the relevant definitions and theorems.

\begin{definition}
        The polynomial $\phi$ is a von Neumann polynomial if all its roots, $r_q$, satisfy $\abs{r_q}\leq 1$.
\end{definition}
\begin{definition}
	The polynomial $\phi$ is a simple von Neumann polynomial is $\phi$ is a von Neumann polynomial and its roots on the unit circle are simple roots.
\end{definition}
\begin{definition}
	For any polynomial $\phi(\xi) = \sum^n_{j=0} a_j\xi^j$, we define the polynomial $\phi^*$ by $\phi^*(\xi) = \sum^n_{j=0} a^*_{n-j} \xi^j$, where $^*$ on the coefficient denotes the complex conjugate.
\end{definition}
\begin{definition}
	For any polynomial $\phi_n(\xi) = \sum^n_{j=0} a_j\xi^j$, we define recursively the polynomial $\phi_{n-1}$ by
	\begin{align}
	\phi_{n-1}(\xi) = \frac{\phi_d^*(0)\phi_d(\xi) - \phi_d(0)\phi_d^*(\xi)}{\xi}.
	\end{align} 
\label{defn:recurse}
\end{definition}
\begin{definition}
	$\phi$ is a Schur polynomial if all its roots, $r_q$, satisfy $\abs{r_q} < 1$.
\end{definition}
\begin{theorem}
	$\phi_n$ is a simple von Neumann polynomial if and only if either 
	\begin{enumerate}[label=(\alph{*})]
		\item $\abs{\phi_n(0)} < \abs{\phi^*_n(0)}$ and $\phi_{n-1}$ is a simple von Neumann polynomial or
		
		\item $\phi_{n-1}$ is identically zero and $\phi'_{n}$ is a Schur polynomial.
	\end{enumerate}
\label{thm:simple vN}
\end{theorem}

\subsection{Amplification polynomials of second order IMEX LMMs}
We start by applying CNAB \eqref{cnab} to the modified test equation \eqref{mte}. Combined with the anzatz $u^n = \xi^n$ and setting $z=\lambda\Delta t$, we get the amplification polynomial
\begin{align}
\Phi_2(\xi) 
= \left(1 - \frac{1}{2}zp\right)\xi^2
- \left(1 + z\left(\frac{3}{2}-p \right)\right)\xi + \frac{1}{2}z(1-p).
\label{amp_cnab}
\end{align}
The next series of steps will show that for $z<0$ (i.e. $\lambda < 0$,) \eqref{amp_cnab} is a simple von Neumann polynomial if and only if $p\geq 1$. This process involves showing that Theorem \ref{thm:simple vN} holds for $\Phi_2$.

We first give $\Phi_2^*$ and $\Phi_1$ as defined by the recursive process in Definition \ref{defn:recurse},
\begin{align}
\Phi_2^*(\xi) = \frac{1}{2}z(1-p) \xi^2 
- \left(1 + z\left(\frac{3}{2} - p \right)\right) \xi
+ 1 - \frac{1}{2}zp, 
\end{align}
and
\begin{align} 
\Phi_1(\xi) = \left(1 - zp + z^2 \left( \frac{1}{2} p - \frac{1}{4} \right) \right)\xi -1 - z(1-p) + z^2 \left( \frac{3}{4} - \frac{1}{2}p \right).
\end{align}
Next, we verify that if $p\geq 1$, then $\abs{\Phi_2(0)} < \abs{\Phi^*_2(0)}$. Reformulating the expression as
\begin{align*}
\abs{\Phi_2(0)} < \abs{\Phi^*_2(0)} 
\iff 
0 < \left(\Phi^*_2(0) \right)^2 - \Phi_2(0)^2
= 1 - zp - \frac{1}{2}z^2\left( \frac{1}{2} - p \right),
\end{align*}
(and keeping in mind that we ask this inequality to hold only for $z<0$,) we find that the contribution from each term  in the rightmost quadratic is positive, thus establishing the claim.
Finally, we show that $\Phi_1$ is simple von Neumann directly. Denoting the root of $\Phi_1$ as $\xi_1$, 
\begin{align*}
	\abs{\xi_1} < 1 
\iff \abs{\frac{(2p-3)z-2}{(2p-1)z-2}} < 1
\iff 0< 8z((p-1)z-1),  
\end{align*}
which holds for all $z<0$ if and only if $p\geq 1$.

Other IMEX LMMs are analyzed in the same way. Their amplification polynomials are recorded for reference in Table \ref{table:amp poly 2}, along with the parameter range for which we observe unconditional stability.

\begin{table}[htb!]
	\centering
	\caption{Amplification polynomial and choice of parameter when applying second order IMEX LMMs to the modified test equation for unconditional stability.}
	\begin{tabular}{lll}
		\toprule[1.5pt] 
		\head{Method} 
		& \head{Amplification Polynomial}
		& \head{$p\in$}
		\\	\midrule 
		CNAB 
		& $\left(1 - \frac{1}{2}zp\right)\xi^2
		- \left(1 + z\left(\frac{3}{2}-p \right)\right)\xi + \frac{1}{2}z(1-p)$
		& $[1,\infty)$
		\\[2.6pt]
		mCNAB 
		& $\left(1 - \frac{9}{16}zp \right) \xi^2 - \left(1 + z\left(\frac{3}{2} - \frac{9}{8} p \right) \right)\xi
		+ \frac{1}{2}z\left(1 - \frac{9}{8}p \right) $
		& $[8/9,\infty)$
		\\[2.6pt]
		CNLF 
		& $\left(1-pz\right) \xi^2 -2z(1-p)\xi -(1+pz)$
		& $[1/2,\infty)$
		\\[2.6pt]
		SBDF2 
		& $\left(\frac{3}{2} - zp\right) \xi^2
		- 2\left(1 + z(1-p)\right) \xi 
		+ \frac{1}{2} + z(1-p)
		$
		& $[3/4,\infty)$
		\\ \bottomrule[1.5pt]
	\end{tabular}
\label{table:amp poly 2}
\end{table}

\subsection{Amplification polynomials of third and fourth oder IMEX LMMs}
We continue with this type of analysis for higher order IMEX LMMs. Because the expressions and manipulations quickly become tedious and cumbersome for higher order methods, the computer algebra system, \textsc{Maple}, was used to for the majority of the calculations. Again, amplification polynomials and parameter ranges are derived. 

\begin{table}[htb!]
	\centering
	\caption{Amplification polynomial and choice of parameter when applying high order IMEX LMMs to the modified test equation for unconditional stability.}
	\begin{tabular}{lll}
		\toprule[1.5pt] 
		\head{Method} 
		& \head{Amplification Polynomial}
		& \head{$p\in$}
		\\	\midrule 
		SBDF3
		& $\left(\frac{11}{6} - zp \right)\xi^3
		- 3\left(1 + z(1-p) \right) \xi^2 
		+ \frac{3}{2}\left(1 + 2z(1-p) \right) \xi 
		$
		& $[7/8, 2]$
		\\
		& \phantom{$\left(\frac{11}{6} - zp \right)\xi^3
			- 3\left(1 + z(1-p) \right) \xi^2$}$- \frac{1}{3}\left(1 +  3z(1-p) \right)$
		\\ [2.6pt]
		SBDF4
		& $\left(\frac{25}{12} - zp \right) \xi^4
		- 4\left(1 + z(1-p) \right)\xi^3
		+ 3\left(1 + 2z(1-p) \right)\xi^2 
		$
		& $[15/16, 5/4]$ 
		\\
		& \phantom{$\left(\frac{25}{12} - zp \right) \xi^4$}$- \frac{4}{3}\left(1 + 3z(1-p) \right) \xi
		+ \frac{1}{4}\left(1 + 4z(1-p)\right)$
		\\ \bottomrule[1.5pt]
	\end{tabular}
	\label{table:amp poly 34}
\end{table}